---
title: "01-SpecimenData"
author: "Galina M. Jönsson"
date: "25/03/2020"
output: html_document
---

Read all NHM specimen data with error radius <10 km and day-level record precision
```{r setup, include=FALSE}
NHMdat <- read.csv("../data/cleanNHMdat/nhm_all_clean_specimens.csv")

# Find the julan day for collection dates
NHMdat$date <- as.Date(NHMdat$date, "%Y-%m-%d")
NHMdat$jDate <- format(NHMdat$date, "%j")
NHMdat$jDate <- as.numeric(NHMdat$jDate)

# Remove any 


### loop through species and find mean and standard deviation for jDate across years
# Create species names vecor
spp <- as.vector(levels(NHMdat$species))

# Create dataframe to populate
jDateSummary <- data.frame(species = as.factor(spp),
                              mean_jDate = as.numeric(NA),
                              SD_jDate = as.numeric(NA))

for(i in 1:length(spp)){
  X <- split(NHMdat, NHMdat$species)
  temp <- X[[i]]
  
  jDateSummary[i,2] <- mean(temp$jDate)
  jDateSummary[i,3] <- sd(temp$jDate)
}
jDateSummary$Min_jDate <- jDateSummary$mean_jDate - 3*jDateSummary$SD_jDate
jDateSummary$Max_jDate <- jDateSummary$mean_jDate + 3*jDateSummary$SD_jDate


### Loop though each species and remove all specimens more than 3 standard deviation form mean jDate

# Split the dat frame into one dtaframe per species 
NHMdatSplit <- split(NHMdat, NHMdat$species)

for(i in 1:length(NHMdatSplit)){
  NHMdatSplit[[i]] <- NHMdatSplit[[i]][(NHMdatSplit[[i]]$jDate < jDateSummary[i,"Max_jDate"]),]
  NHMdatSplit[[i]] <- NHMdatSplit[[i]][(NHMdatSplit[[i]]$jDate > jDateSummary[i,"Min_jDate"]),]
}


### Loop though each species, find the numebr of records per year and merge this data with as a separate column

# Split the dat frame into one dtaframe per species 
NHMdatSplit <- split(NHMdat, NHMdat$species)

for(i in 1:length(NHMdatSplit)){
  require(dplyr)
  SpecimepnPerSpp <- NHMdatSplit[[i]] %>% group_by(year) %>% tally()
  NHMdatSplit[[i]] <- merge(NHMdatSplit[[i]], SpecimepnPerSpp, by="year", all.x = TRUE)
  NHMdatSplit[[i]] <- NHMdatSplit[[i]][(NHMdatSplit[[i]]$n > 4),]
}

# Combine into one data frame
NHMclean <- bind_rows(NHMdatSplit, .id = "column_label")

# Find records per species
require(dplyr)
SpecimepnPerSpp <- NHMclean %>% group_by(species) %>% tally()
SpecimepnPerSpp
```

### Univoltine species
First, I will subset the univoltine species (designation according to Dennis et al., 2017; http://dx.doi.org/10.1016/j.ecolind.2017.01.009; disregarding Erynnis tages, which is known to be bivoltine in the UK, this is also apparent from histograms)

The univoltine species are:
- Thymelicus lineola  
- Pyronia tithonus  
- Callophrys rubi  
- Pyrgus malvae  
- Ochlodes sylvanus  
- Melanargia galathea  
- Maniola jurtina  
- Anthocharis cardamines  
- Favonius quercus  
- Aphantopus hyperantus  
- Thymelicus sylvestris  



```{r temporal-spread-raw}
uniVolSpp <- c("Thymelicus lineola", "Pyronia tithonus", "Callophrys rubi", "Pyrgus malvae", "Ochlodes sylvanus", "Melanargia galathea", "Maniola jurtina", "Anthocharis cardamines", "Favonius quercus", "Aphantopus hyperantus", "Thymelicus sylvestris")
length(uniVolSpp)

NHMcleanUniVol <- subset(NHMclean, species %in% uniVolSpp)
NHMcleanUniVol <- subset(NHMcleanUniVol, year < 1970)
NHMcleanUniVol <- droplevels(NHMcleanUniVol)

for(i in 1:length(uniVolSpp)){
  hist((subset(NHMcleanUniVol, species == uniVolSpp[i]))$jDate, xlab = "Julian day", ylab = "Specimen records", main = uniVolSpp[i])
}

# Find records per species
require(dplyr)
SpecimepnPerSpp <- NHMcleanUniVol %>% group_by(species) %>% tally()
SpecimepnPerSpp

```





#### Maniola jurtina (Meadow brown)
```{r maniola-jurtina}
Maniola_jurtina <- subset(NHMcleanUniVol, species == "Maniola jurtina")
Maniola_jurtina$year <- as.factor(as.character(Maniola_jurtina$year))

# I will remove four obviou outlier records (February 1947, Feb 1948, Feb 1952 and March 1922)
Maniola_jurtina <- subset(Maniola_jurtina, month > 3)

# Plot UK outline
plot_GIS(UK, new.window = FALSE)
# Add these grid references to an exisiting plot
plotUK_gr(Maniola_jurtina$km10grid, col="red")

Maniola_jurtinaYearly <- data.frame(year = as.factor(unique(Maniola_jurtina$year)),
                                  percentile10 = as.numeric(NA),
                                  percentile90 = as.numeric(NA),
                                  median = as.numeric(NA),
                                  earliest = as.numeric(NA),
                                  CollectionPeriod = as.numeric(NA))

for(i in 1:length(unique(Maniola_jurtina$year))){
  temp <- subset(Maniola_jurtina, year == (unique(Maniola_jurtina$year)[i]))
  Maniola_jurtinaYearly[i,2] <- quantile((temp$jDate), 0.10)
  Maniola_jurtinaYearly[i,3] <- quantile((temp$jDate), 0.90)  
  Maniola_jurtinaYearly[i,4] <- median(temp$jDate) 
  Maniola_jurtinaYearly[i,5] <- min(temp$jDate) 
  Maniola_jurtinaYearly[i,6] <- Erynnis_tagesYearly[i,3] - Erynnis_tagesYearly[i,2]
}

# Merge Maniola_jurtinaYearly with the Maniola_jurtina by year
Maniola_jurtina <- merge(Maniola_jurtina, Maniola_jurtinaYearly, by="year", all.x=TRUE)

hist(Maniola_jurtina$jDate, xlab = "Julian day", ylab = "Specimen records", main = "Maniola jurtina")

Maniola_jurtinaYearly$year <- as.numeric(as.character(Maniola_jurtinaYearly$year))

require(ggplot2)
ggplot(Maniola_jurtinaYearly, aes(x=year, y=percentile10)) +
  geom_point() + 
  geom_smooth(method=lm)
```

#### Pyronia tithonus (Gatekeeper)
```{r pyronia-tithonus}
Pyronia_tithonus <- subset(NHMcleanUniVol, species == "Pyronia tithonus")
Pyronia_tithonus$year <- as.factor(as.character(Pyronia_tithonus$year))

# I will remove five obvious outlier records (January 1942, Two records from April 1928, two records from November 1902)
Pyronia_tithonus <- subset(Pyronia_tithonus, month > 4)
Pyronia_tithonus <- subset(Pyronia_tithonus, month < 11)

# Plot UK outline
plot_GIS(UK, new.window = FALSE)
# Add these grid references to an exisiting plot
plotUK_gr(Pyronia_tithonus$km10grid, col="red")

Pyronia_tithonusYearly <- data.frame(year = as.factor(unique(Pyronia_tithonus$year)),
                                  percentile10 = as.numeric(NA),
                                  percentile90 = as.numeric(NA),
                                  median = as.numeric(NA),
                                  earliest = as.numeric(NA),
                                  CollectionPeriod = as.numeric(NA))

for(i in 1:length(unique(Pyronia_tithonus$year))){
  temp <- subset(Pyronia_tithonus, year == (unique(Pyronia_tithonus$year)[i]))
  Pyronia_tithonusYearly[i,2] <- quantile((temp$jDate), 0.10)
  Pyronia_tithonusYearly[i,3] <- quantile((temp$jDate), 0.90)  
  Pyronia_tithonusYearly[i,4] <- median(temp$jDate) 
  Pyronia_tithonusYearly[i,5] <- min(temp$jDate) 
  Pyronia_tithonusYearly[i,6] <- Pyronia_tithonusYearly[i,3] - Pyronia_tithonusYearly[i,2]
}

# Merge Pyronia_tithonusYearly with the Pyronia_tithonus by year
Pyronia_tithonus <- merge(Pyronia_tithonus, Pyronia_tithonusYearly, by="year", all.x=TRUE)



Pyronia_tithonusYearly$year <- as.numeric(as.character(Pyronia_tithonusYearly$year))

hist(Pyronia_tithonus$jDate, xlab = "Julian day", ylab = "Specimen records", main = "Pyronia tithonus")

require(ggplot2)
ggplot(Pyronia_tithonusYearly, aes(x=year, y=percentile10)) +
  geom_point() + 
  geom_smooth(method=lm)
```



#### Callophrys rubi (Green hairstreak)
```{r Callophrys-rubi}

Callophrys_rubi <- subset(NHMcleanUniVol, species == "Callophrys rubi")
Callophrys_rubi$year <- as.factor(as.character(Callophrys_rubi$year))

# I will remove all 1913 records (15 records in total), which are obvious outliers as most dates are February or March and if these are removed there are not enogh (five or more) sensical records
Callophrys_rubi <- subset(Callophrys_rubi, !year == "1913")

# Plot UK outline
plot_GIS(UK, new.window = FALSE)
# Add these grid references to an exisiting plot
plotUK_gr(Callophrys_rubi$km10grid, col="red")

Callophrys_rubiYearly <- data.frame(year = as.factor(unique(Callophrys_rubi$year)),
                                  percentile10 = as.numeric(NA),
                                  percentile90 = as.numeric(NA),
                                  median = as.numeric(NA),
                                  earliest = as.numeric(NA),
                                  CollectionPeriod = as.numeric(NA))

for(i in 1:length(unique(Callophrys_rubi$year))){
  temp <- subset(Callophrys_rubi, year == (unique(Callophrys_rubi$year)[i]))
  Callophrys_rubiYearly[i,2] <- quantile((temp$jDate), 0.10)
  Callophrys_rubiYearly[i,3] <- quantile((temp$jDate), 0.90)  
  Callophrys_rubiYearly[i,4] <- median(temp$jDate) 
  Callophrys_rubiYearly[i,5] <- min(temp$jDate) 
  Callophrys_rubiYearly[i,6] <- Callophrys_rubiYearly[i,3] - Callophrys_rubiYearly[i,2]
}

# Merge Callophrys_rubiYearly with the Callophrys_rubi by year
Callophrys_rubi <- merge(Callophrys_rubi, Callophrys_rubiYearly, by="year", all.x=TRUE)

hist(Callophrys_rubi$jDate, xlab = "Julian day", ylab = "Specimen records", main = "Callophrys rubi")

Callophrys_rubiYearly$year <- as.numeric(as.character(Callophrys_rubiYearly$year))

require(ggplot2)
ggplot(Callophrys_rubiYearly, aes(x=year, y=percentile10)) +
  geom_point() + 
  geom_smooth(method=lm)

```


#### Pyrgus malvae (Grizzled skipper)
```{r Pyrgus-malvae}
Pyrgus_malvae <- subset(NHMcleanUniVol, species == "Pyrgus malvae")
Pyrgus_malvae$year <- as.factor(as.character(Pyrgus_malvae$year))

# I will remove all records belonging to the very rare second generation, all records with jdat > 210
Pyrgus_malvae <- subset(Pyrgus_malvae, jDate < 210)

# Plot UK outline
plot_GIS(UK, new.window = FALSE)
# Add these grid references to an exisiting plot
plotUK_gr(Pyrgus_malvae$km10grid, col="red")

Pyrgus_malvaeYearly <- data.frame(year = as.factor(unique(Pyrgus_malvae$year)),
                                  percentile10 = as.numeric(NA),
                                  percentile90 = as.numeric(NA),
                                  median = as.numeric(NA),
                                  earliest = as.numeric(NA),
                                  CollectionPeriod = as.numeric(NA))

for(i in 1:length(unique(Pyrgus_malvae$year))){
  temp <- subset(Pyrgus_malvae, year == (unique(Pyrgus_malvae$year)[i]))
  Pyrgus_malvaeYearly[i,2] <- quantile((temp$jDate), 0.10)
  Pyrgus_malvaeYearly[i,3] <- quantile((temp$jDate), 0.90)  
  Pyrgus_malvaeYearly[i,4] <- median(temp$jDate) 
  Pyrgus_malvaeYearly[i,5] <- min(temp$jDate) 
  Pyrgus_malvaeYearly[i,6] <- Pyrgus_malvaeYearly[i,3] - Pyrgus_malvaeYearly[i,2]
}

# Merge Pyrgus_malvaeYearly with the Pyrgus_malvae by year
Pyrgus_malvae <- merge(Pyrgus_malvae, Pyrgus_malvaeYearly, by="year", all.x=TRUE)

hist(Pyrgus_malvae$jDate, xlab = "Julian day", ylab = "Specimen records", main = "Pyrgus malvae")


Pyrgus_malvaeYearly$year <- as.numeric(as.character(Pyrgus_malvaeYearly$year))


require(ggplot2)
ggplot(Pyrgus_malvaeYearly, aes(x=year, y=percentile10)) +
  geom_point() + 
  geom_smooth(method=lm)

```



#### Ochlodes sylvanus (Large skipper)
```{r Ochlodes-sylvanus}
Ochlodes_sylvanus <- subset(NHMcleanUniVol, species == "Ochlodes sylvanus")
Ochlodes_sylvanus$year <- as.factor(as.character(Ochlodes_sylvanus$year))

# Plot UK outline
plot_GIS(UK, new.window = FALSE)
# Add these grid references to an exisiting plot
plotUK_gr(Ochlodes_sylvanus$km10grid, col="red")

Ochlodes_sylvanusYearly <- data.frame(year = as.factor(unique(Ochlodes_sylvanus$year)),
                                  percentile10 = as.numeric(NA),
                                  percentile90 = as.numeric(NA),
                                  median = as.numeric(NA),
                                  earliest = as.numeric(NA),
                                  CollectionPeriod = as.numeric(NA))

for(i in 1:length(unique(Ochlodes_sylvanus$year))){
  temp <- subset(Ochlodes_sylvanus, year == (unique(Ochlodes_sylvanus$year)[i]))
  Ochlodes_sylvanusYearly[i,2] <- quantile((temp$jDate), 0.10)
  Ochlodes_sylvanusYearly[i,3] <- quantile((temp$jDate), 0.90)  
  Ochlodes_sylvanusYearly[i,4] <- median(temp$jDate) 
  Ochlodes_sylvanusYearly[i,5] <- min(temp$jDate) 
  Ochlodes_sylvanusYearly[i,6] <- Ochlodes_sylvanusYearly[i,3] - Ochlodes_sylvanusYearly[i,2]
}

# Merge Ochlodes_sylvanusYearly with the Ochlodes_sylvanus by year
Ochlodes_sylvanus <- merge(Ochlodes_sylvanus, Ochlodes_sylvanusYearly, by="year", all.x=TRUE)

hist(Ochlodes_sylvanus$jDate, xlab = "Julian day", ylab = "Specimen records", main = "Ochlodes sylvanus")


Ochlodes_sylvanusYearly$year <- as.numeric(as.character(Ochlodes_sylvanusYearly$year))

require(ggplot2)
ggplot(Ochlodes_sylvanusYearly, aes(x=year, y=percentile10)) +
  geom_point() + 
  geom_smooth(method=lm)

```



#### Melanargia galathea (Marbled white)
```{r Melanargia-galathea}
Melanargia_galathea <- subset(NHMcleanUniVol, species == "Melanargia galathea")
Melanargia_galathea$year <- as.factor(as.character(Melanargia_galathea$year))

# I will remove two obvious outliers (February 1948 and October 1921)
Melanargia_galathea <- subset(Melanargia_galathea, month < 10)
Melanargia_galathea <- subset(Melanargia_galathea, month > 2)

# Plot UK outline
plot_GIS(UK, new.window = FALSE)
# Add these grid references to an exisiting plot
plotUK_gr(Melanargia_galathea$km10grid, col="red")

Melanargia_galatheaYearly <- data.frame(year = as.factor(unique(Melanargia_galathea$year)),
                                  percentile10 = as.numeric(NA),
                                  percentile90 = as.numeric(NA),
                                  median = as.numeric(NA),
                                  earliest = as.numeric(NA),
                                  CollectionPeriod = as.numeric(NA))

for(i in 1:length(unique(Melanargia_galathea$year))){
  temp <- subset(Melanargia_galathea, year == (unique(Melanargia_galathea$year)[i]))
  Melanargia_galatheaYearly[i,2] <- quantile((temp$jDate), 0.10)
  Melanargia_galatheaYearly[i,3] <- quantile((temp$jDate), 0.90)  
  Melanargia_galatheaYearly[i,4] <- median(temp$jDate) 
  Melanargia_galatheaYearly[i,5] <- min(temp$jDate) 
  Melanargia_galatheaYearly[i,6] <- Melanargia_galatheaYearly[i,3] - Melanargia_galatheaYearly[i,2]
}

# Merge Melanargia_galatheaYearly with the Melanargia_galathea by year
Melanargia_galathea <- merge(Melanargia_galathea, Melanargia_galatheaYearly, by="year", all.x=TRUE)

hist(Melanargia_galathea$jDate, xlab = "Julian day", ylab = "Specimen records", main = "Melanargia galathea")


Melanargia_galatheaYearly$year <- as.numeric(as.character(Melanargia_galatheaYearly$year))

require(ggplot2)
ggplot(Melanargia_galatheaYearly, aes(x=year, y=percentile10)) +
  geom_point() + 
  geom_smooth(method=lm)

```

 

#### Anthocharis cardamines (Orange-tip)
```{r Anthocharis-cardamines}
Anthocharis_cardamines <- subset(NHMcleanUniVol, species == "Anthocharis cardamines")
Anthocharis_cardamines$year <- as.factor(as.character(Anthocharis_cardamines$year))


# I will remove one obvious outlier (February 1911)
Anthocharis_cardamines <- subset(Anthocharis_cardamines, month > 2)

# Plot UK outline
plot_GIS(UK, new.window = FALSE)
# Add these grid references to an exisiting plot
plotUK_gr(Anthocharis_cardamines$km10grid, col="red")

Anthocharis_cardaminesYearly <- data.frame(year = as.factor(unique(Anthocharis_cardamines$year)),
                                  percentile10 = as.numeric(NA),
                                  percentile90 = as.numeric(NA),
                                  median = as.numeric(NA),
                                  earliest = as.numeric(NA),
                                  CollectionPeriod = as.numeric(NA))

for(i in 1:length(unique(Anthocharis_cardamines$year))){
  temp <- subset(Anthocharis_cardamines, year == (unique(Anthocharis_cardamines$year)[i]))
  Anthocharis_cardaminesYearly[i,2] <- quantile((temp$jDate), 0.10)
  Anthocharis_cardaminesYearly[i,3] <- quantile((temp$jDate), 0.90)  
  Anthocharis_cardaminesYearly[i,4] <- median(temp$jDate) 
  Anthocharis_cardaminesYearly[i,5] <- min(temp$jDate) 
  Anthocharis_cardaminesYearly[i,6] <- Anthocharis_cardaminesYearly[i,3] - Anthocharis_cardaminesYearly[i,2]
}

# Merge Anthocharis_cardaminesYearly with the Anthocharis_cardamines by year
Anthocharis_cardamines <- merge(Anthocharis_cardamines, Anthocharis_cardaminesYearly, by="year", all.x=TRUE)

hist(Anthocharis_cardamines$jDate, xlab = "Julian day", ylab = "Specimen records", main = "Anthocharis cardamines")


Anthocharis_cardaminesYearly$year <- as.numeric(as.character(Anthocharis_cardaminesYearly$year))

require(ggplot2)
ggplot(Anthocharis_cardaminesYearly, aes(x=year, y=percentile10)) +
  geom_point() + 
  geom_smooth(method=lm)
```



#### Favonius quercus (Purple hairstreak)
```{r Favonius-quercus}
Favonius_quercus <- subset(NHMcleanUniVol, species == "Favonius quercus")
Favonius_quercus$year <- as.factor(as.character(Favonius_quercus$year))

# Plot UK outline
plot_GIS(UK, new.window = FALSE)
# Add these grid references to an exisiting plot
plotUK_gr(Favonius_quercus$km10grid, col="red")

Favonius_quercusYearly <- data.frame(year = as.factor(unique(Favonius_quercus$year)),
                                  percentile10 = as.numeric(NA),
                                  percentile90 = as.numeric(NA),
                                  median = as.numeric(NA),
                                  earliest = as.numeric(NA),
                                  CollectionPeriod = as.numeric(NA))

for(i in 1:length(unique(Favonius_quercus$year))){
  temp <- subset(Favonius_quercus, year == (unique(Favonius_quercus$year)[i]))
  Favonius_quercusYearly[i,2] <- quantile((temp$jDate), 0.10)
  Favonius_quercusYearly[i,3] <- quantile((temp$jDate), 0.90)  
  Favonius_quercusYearly[i,4] <- median(temp$jDate) 
  Favonius_quercusYearly[i,5] <- min(temp$jDate) 
  Favonius_quercusYearly[i,6] <- Favonius_quercusYearly[i,3] - Favonius_quercusYearly[i,2]
}

# Merge Favonius_quercusYearly with the Favonius_quercus by year
Favonius_quercus <- merge(Favonius_quercus, Favonius_quercusYearly, by="year", all.x=TRUE)

hist(Favonius_quercus$jDate, xlab = "Julian day", ylab = "Specimen records", main = "Favonius quercus")

Favonius_quercusYearly$year <- as.numeric(as.character(Favonius_quercusYearly$year))

require(ggplot2)
ggplot(Favonius_quercusYearly, aes(x=year, y=percentile10)) +
  geom_point() + 
  geom_smooth(method=lm)
```



#### Aphantopus hyperantus (Ringlet)
```{r Aphantopus-hyperantus}
Aphantopus_hyperantus <- subset(NHMcleanUniVol, species == "Aphantopus hyperantus")
Aphantopus_hyperantus$year <- as.factor(as.character(Aphantopus_hyperantus$year))


# I will remove four obvious outliers (January 1924, February 1948, March 1943 amd October 1907)
Aphantopus_hyperantus <- subset(Aphantopus_hyperantus, month > 3)
Aphantopus_hyperantus <- subset(Aphantopus_hyperantus, month < 10)

# Plot UK outline
plot_GIS(UK, new.window = FALSE)
# Add these grid references to an exisiting plot
plotUK_gr(Aphantopus_hyperantus$km10grid, col="red")

Aphantopus_hyperantusYearly <- data.frame(year = as.factor(unique(Aphantopus_hyperantus$year)),
                                  percentile10 = as.numeric(NA),
                                  percentile90 = as.numeric(NA),
                                  median = as.numeric(NA),
                                  earliest = as.numeric(NA),
                                  CollectionPeriod = as.numeric(NA))

for(i in 1:length(unique(Aphantopus_hyperantus$year))){
  temp <- subset(Aphantopus_hyperantus, year == (unique(Aphantopus_hyperantus$year)[i]))
  Aphantopus_hyperantusYearly[i,2] <- quantile((temp$jDate), 0.10)
  Aphantopus_hyperantusYearly[i,3] <- quantile((temp$jDate), 0.90)  
  Aphantopus_hyperantusYearly[i,4] <- median(temp$jDate) 
  Aphantopus_hyperantusYearly[i,5] <- min(temp$jDate) 
  Aphantopus_hyperantusYearly[i,6] <- Aphantopus_hyperantusYearly[i,3] - Aphantopus_hyperantusYearly[i,2]
}

# Merge Aphantopus_hyperantusYearly with the Aphantopus_hyperantus by year
Aphantopus_hyperantus <- merge(Aphantopus_hyperantus, Aphantopus_hyperantusYearly, by="year", all.x=TRUE)

hist(Aphantopus_hyperantus$jDate, xlab = "Julian day", ylab = "Specimen records", main = "Aphantopus hyperantus")

Aphantopus_hyperantusYearly$year <- as.numeric(as.character(Aphantopus_hyperantusYearly$year))

require(ggplot2)
ggplot(Aphantopus_hyperantusYearly, aes(x=year, y=percentile10)) +
  geom_point() + 
  geom_smooth(method=lm)

```



#### Thymelicus sylvestris (Small skipper)
```{r Thymelicus-sylvestris}
Thymelicus_sylvestris <- subset(NHMcleanUniVol, species == "Thymelicus sylvestris")
Thymelicus_sylvestris$year <- as.factor(as.character(Thymelicus_sylvestris$year))

# I will remove one obvious outlier (November 1947)
Thymelicus_sylvestris <- subset(Thymelicus_sylvestris, month < 11)

# Plot UK outline
plot_GIS(UK, new.window = FALSE)
# Add these grid references to an exisiting plot
plotUK_gr(Thymelicus_sylvestris$km10grid, col="red")

Thymelicus_sylvestrisYearly <- data.frame(year = as.factor(unique(Thymelicus_sylvestris$year)),
                                  percentile10 = as.numeric(NA),
                                  percentile90 = as.numeric(NA),
                                  median = as.numeric(NA),
                                  earliest = as.numeric(NA),
                                  CollectionPeriod = as.numeric(NA))

for(i in 1:length(unique(Thymelicus_sylvestris$year))){
  temp <- subset(Thymelicus_sylvestris, year == (unique(Thymelicus_sylvestris$year)[i]))
  Thymelicus_sylvestrisYearly[i,2] <- quantile((temp$jDate), 0.10)
  Thymelicus_sylvestrisYearly[i,3] <- quantile((temp$jDate), 0.90)  
  Thymelicus_sylvestrisYearly[i,4] <- median(temp$jDate) 
  Thymelicus_sylvestrisYearly[i,5] <- min(temp$jDate) 
  Thymelicus_sylvestrisYearly[i,6] <- Thymelicus_sylvestrisYearly[i,3] - Thymelicus_sylvestrisYearly[i,2]
}

# Merge Thymelicus_sylvestrisYearly with the Thymelicus_sylvestris by year
Thymelicus_sylvestris <- merge(Thymelicus_sylvestris, Thymelicus_sylvestrisYearly, by="year", all.x=TRUE)

hist(Thymelicus_sylvestris$jDate, xlab = "Julian day", ylab = "Specimen records", main = "Thymelicus sylvestris")

Thymelicus_sylvestrisYearly$year <- as.numeric(as.character(Thymelicus_sylvestrisYearly$year))

require(ggplot2)
ggplot(Thymelicus_sylvestrisYearly, aes(x=year, y=percentile10)) +
  geom_point() + 
  geom_smooth(method=lm)
```


#### Thymelicus lineola (Essex skipper)
```{r Thymelicus-lineola}
Thymelicus_lineola <- subset(NHMcleanUniVol, species == "Thymelicus lineola")
Thymelicus_lineola$year <- as.factor(as.character(Thymelicus_lineola$year))

# Plot UK outline
plot_GIS(UK, new.window = FALSE)
# Add these grid references to an exisiting plot
plotUK_gr(Thymelicus_lineola$km10grid, col="red")

Thymelicus_lineolaYearly <- data.frame(year = as.factor(unique(Thymelicus_lineola$year)),
                                  percentile10 = as.numeric(NA),
                                  percentile90 = as.numeric(NA),
                                  median = as.numeric(NA),
                                  earliest = as.numeric(NA),
                                  CollectionPeriod = as.numeric(NA))

for(i in 1:length(unique(Thymelicus_lineola$year))){
  temp <- subset(Thymelicus_lineola, year == (unique(Thymelicus_lineola$year)[i]))
  Thymelicus_lineolaYearly[i,2] <- quantile((temp$jDate), 0.10)
  Thymelicus_lineolaYearly[i,3] <- quantile((temp$jDate), 0.90)  
  Thymelicus_lineolaYearly[i,4] <- median(temp$jDate) 
  Thymelicus_lineolaYearly[i,5] <- min(temp$jDate) 
  Thymelicus_lineolaYearly[i,6] <- Thymelicus_lineolaYearly[i,3] - Thymelicus_lineolaYearly[i,2]
}

# Merge Thymelicus_lineolaYearly with the Thymelicus_lineola by year
Thymelicus_lineola <- merge(Thymelicus_lineola, Thymelicus_lineolaYearly, by="year", all.x=TRUE)

hist(Thymelicus_lineola$jDate, xlab = "Julian day", ylab = "Specimen records", main = "Thymelicus lineola")

Thymelicus_lineolaYearly$year <- as.numeric(as.character(Thymelicus_lineolaYearly$year))

require(ggplot2)
ggplot(Thymelicus_lineolaYearly, aes(x=year, y=percentile10)) +
  geom_point() + 
  geom_smooth(method=lm)
```









### Bivoltine species

#### Erynnis tages (dingy skipper)
```{r erynnis-tages}
Erynnis_tages <- subset(NHMcleanUniVol, species == "Erynnis tages") # 511 obs
Erynnis_tages$year <- as.factor(as.character(Erynnis_tages$year))

## Map records
require(BRCmap)
#Load UK data
data(UK)

# Plot UK outline
plot_GIS(UK, new.window = FALSE)
  
# Add these grid references to an exisiting plot
plotUK_gr(Erynnis_tages$km10grid, col="red")

Erynnis_tagesYearly <- data.frame(year = as.factor(unique(Erynnis_tages$year)),
                                  percentile10 = as.numeric(NA),
                                  percentile90 = as.numeric(NA),
                                  median = as.numeric(NA),
                                  earliest = as.numeric(NA),
                                  CollectionPeriod = as.numeric(NA))

for(i in 1:length(unique(Erynnis_tages$year))){
  temp <- subset(Erynnis_tages, year == (unique(Erynnis_tages$year)[i]))
  Erynnis_tagesYearly[i,2] <- quantile((temp$jDate), 0.10)
  Erynnis_tagesYearly[i,3] <- quantile((temp$jDate), 0.90)  
  Erynnis_tagesYearly[i,4] <- median(temp$jDate) 
  Erynnis_tagesYearly[i,5] <- min(temp$jDate) 
  Erynnis_tagesYearly[i,6] <- Erynnis_tagesYearly[i,3] - Erynnis_tagesYearly[i,2]
}


# Merge Erynnis_tagesYearly with the Erynnis_tages by year
Erynnis_tages <- merge(Erynnis_tages, Erynnis_tagesYearly, by="year", all.x=TRUE)


Erynnis_tagesYearly$year <- as.numeric(as.character(Erynnis_tagesYearly$year))

require(ggplot2)
ggplot(Erynnis_tagesYearly, aes(x=year, y=percentile10)) +
  geom_point() + 
  geom_smooth(method=lm)

ggplot(Erynnis_tagesYearly, aes(x=year, y=earliest)) +
  geom_point() + 
  geom_smooth(method=lm)
```
