---
title: "01-SpecimenData"
author: "Galina M. Jönsson"
date: "25/03/2020"
output: html_document
---

Read all NHM specimen data with error radius <10 km and day-level record precision
```{r setup, include=FALSE}
NHMdat <- read.csv("../data/cleanNHMdat/nhm_all_clean_specimens.csv")

# Find records per species
require(dplyr)
SpecimepnPerSpp <- NHMdat %>% group_by(species) %>% tally()
SpecimepnPerSpp

# Find the julan day for collection dates
NHMdat$date <- as.Date(NHMdat$date, "%Y-%m-%d")
NHMdat$jDate <- format(NHMdat$date, "%j")
NHMdat$jDate <- as.numeric(NHMdat$jDate)


### loop through species and find mean and standard deviation for jDate across years
# Create species names vecor
spp <- as.vector(levels(NHMdat$species))

# Create dataframe to populate
jDateSummary <- data.frame(species = as.factor(spp),
                              mean_jDate = as.numeric(NA),
                              SD_jDate = as.numeric(NA))

for(i in 1:length(spp)){
  X <- split(NHMdat, NHMdat$species)
  temp <- X[[i]]
  
  jDateSummary[i,2] <- mean(temp$jDate)
  jDateSummary[i,3] <- sd(temp$jDate)
}
jDateSummary$ThreeSD_jDate <- jDateSummary$SD_jDate*3



### loop through species and find remove specimens collected more or less than three standard deviations from mean jDate across years
spp <- as.vector(levels(NHMdat$species))

# Plit NMdat into multiple dataframes, one per species
NHMdatSplit <- split(NHMdat, NHMdat$species)

# loop through
for(i in 1:length(spp)){
  temp <- NHMdatSplit[[i]]
  for(j in 1:nrow(temp))
    temp$jDate[j,]
  if(temp$jDate[j,] < (jDateSummary))
  
  jDateSummary[i,2] <- mean(temp$jDate)
  jDateSummary[i,3] <- sd(temp$jDate)
}
jDateSummary$ThreeSD_jDate <- jDateSummary$SD_jDate*3

jDateSummary
```


```{r temporal-spread-raw}
require(ggplot2)
p <- ggplot((NHMdat %>%
    group_by(species, year) %>%
    tally()), aes(year, n)) + geom_point() +
    facet_grid(rows = vars(species))
p
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
